Predicting trends in the quality of state-of-the-art neural networks without access to training or testing data

### 2020-07-02 submission/background.tex
minor edits

There is also a large body of work on using norm-based metrics to bound generalization error~\cite{NTS15, BFT17_TR, LMBx18_TR}.
In this area, theoretical work aims to prove generalization bounds, and applied work uses these norms to construct regularizers to improve training.
While we do find that norms provide relatively good quality metrics, at least for distinguishing among well-trained versus more-well-trained models, we are not interested in proving generalization bounds or developing new regularizers.

>>>>

There is also a large body of work on using norm-based metrics to bound generalization error~\cite{NTS15, BFT17_TR, LMBx18_TR}.
In this area, theoretical work aims to prove generalization bounds, and applied work uses these norms to construct regularizers to improve training.
While we do find that norms provide relatively good quality metrics, at least for distinguishing among \nred{(moderatly)} well-trained versus 
\nred{very-well-trained} models, \nred{of increasing deopth an accuracy}, 
we are not interested in proving generalization bounds or developing new regularizers.

### 2020-07-02 submission/cv_models.tex
minor edits

We have examined the performance of the four quality metrics (Log Frobenius norm, Log Spectral norm, Weighted Alpha, and Log $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without BN), and it plots the four quality metrics versus the reported test accuracies~\cite{pytorch},% 
\footnote{That is, these test accuracies have been previously reported and made publicly-available by others.  We take them as given, and we do not attempt to reproduce/verify them, since we do not permit ourselves any access to training/test data.}
as well as a basic linear regression line. 
All four metrics correlate quite well with the reported Top1 accuracies, with smaller norms and smaller values of $\hat{\alpha}$ implying better generalization (i.e., greater accuracy, lower error). 
While all four metrics perform well, notice that the Log $\alpha$-Norm metric ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$) performs best (with an RMSE of $0.42$, see Table~\ref{table:cv-models}); and the Weighted Alpha metric ($\hat\alpha =\alpha\log\lambda_{max} $), which is an approximation to the Log $\alpha$-Norm metric~\cite{MM20_unpub_work}, performs second best (with an RMSE of $0.48$, see Table~\ref{table:cv-models}).

@@

See Table~\ref{table:cv-models} for a summary of results for Top1 accuracies for all four metrics for the VGG, ResNet, and DenseNet series.
Similar results (not shown) are obtained for the Top5 accuracies.
Overall, for the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and the Weighted Alpha metric performs best.
These model series are all very well-trodden, and thus our results indicate that norm-based metrics and PL-based metrics can both distinguish among a series of well-trained versus more-well-trained models, with PL-based metrics performing somewhat (i.e., quantitatively) better.

>>>>

We have examined the performance of the four quality metrics (Log Frobenius norm, Log Spectral norm, Weighted Alpha, and Log $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without \nred{BatchNormalization, denoted} BN), and it plots the four quality metrics versus the reported test accuracies~\cite{pytorch},% 
\footnote{That is, these test accuracies have been previously reported and made publicly-available by others.  We take them as given, and we do not attempt to reproduce/verify them, since we do not permit ourselves any access to training/test data.}
as well as a basic linear regression line. 
All four metrics correlate quite well with the reported Top1 accuracies, with smaller norms and smaller values of $\hat{\alpha}$ implying better generalization (i.e., greater accuracy, lower error). 
While all four metrics perform well, notice that the Log $\alpha$-Norm metric ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$) performs best (with an RMSE of $0.42$, see Table~\ref{table:cv-models}); and the Weighted Alpha metric ($\hat\alpha =\alpha\log\lambda_{max} $), which is an approximation to the Log $\alpha$-Norm metric~\cite{MM20_unpub_work}, performs second best (with an RMSE of $0.48$, see Table~\ref{table:cv-models}).

@@

See Table~\ref{table:cv-models} for a summary of results for Top1 accuracies for all four metrics for the VGG, ResNet, and DenseNet series.
Similar results (not shown) are obtained for the Top5 accuracies.
Overall, for the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and the Weighted Alpha metric performs best.
These model series are all very well-trodden, and thus our results indicate that norm-based metrics and PL-based metrics can both distinguish among a series of well-trained versus very-well-trained models, with PL-based metrics performing somewhat (i.e., quantitatively) better.

### 2020-07-02 submission/nlp_models.tex
minor edits

The OpenAI GPT and GPT2 models provide us with the opportunity to analyze two effects: training the same model with different data set sizes; and increasing the sizes of both the data set and the architectures simultaneously.
These models have the remarkable ability to generate fake text that appears to the human to be real, and they have generated significant media attention because of the potential for their misuse.
For this reason, the original GPT model released by OpenAI was trained on a deficient data set, rendering the model interesting but not fully functional.  
Later, OpenAI released a much improved model, GPT2-small, which has the same architecture and number of layers as GPT, but which has been trained on a larger and better data set (and with other changes), making it remarkably good at generating (near) human-quality fake text.  
%
By comparing the poorly-trained (i.e., ``bad'') GPT to the well-trained (i.e., ``good'') GPT2-small, we can identify empirical indicators for when a model has in fact been poorly-trained and thus may perform poorly when deployed.
By comparing GPT2-medium to GPT2-large to GPT2-xl, we can examine the effect of increasing data set and model size simultaneously, as well as analyze well-trained versus more-well-trained models.

@@

We now look across series of increasingly improving GPT2 models (i.e., we consider well-trained versus more-well-trained questions), by examining both the PL exponent $\alpha$ as well as the Log Norm metrics.  
In general, as we move from GPT2-medium to GPT2-xl, histograms for both $\alpha$ exponents and the Log Norm metrics downshift from larger to smaller values. 
For example, see Figure \ref{fig:gpt2-histograms}, which shows the histograms over the layer weight matrices for fitted PL exponent ($\alpha$) and the Log Alpha Norm ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$)~metric.

We see that the average $\alpha$ decreases with increasing model size, although the differences are less noticeable between the differing well-trained versus more-well-trained GTP2 models than between the poorly-trained versus well-trained GPT and GPT2-small models.
Unlike GPT, however, the layer Log Alpha Norms behave more as expected for GPT2 layers, with the larger models consistently having smaller norms. 
Similarly, the Log Spectral Norm also decreases on average with the larger models (not shown).  
As expected, the norm metrics can indeed distinguish among well-trained versus more-well-trained models among a series well-trained models.

>>>>

The OpenAI GPT and GPT2 models provide us with the opportunity to analyze two effects: training the same model with different data set sizes; and increasing the sizes of both the data set and the architectures simultaneously.
These models have the remarkable ability to generate fake text that appears to the human to be real, and they have generated significant media attention because of the potential for their misuse.
For this reason, the original GPT model released by OpenAI was trained on a deficient data set, rendering the model interesting but not fully functional.  
Later, OpenAI released a much improved model, GPT2-small, which has the same architecture and number of layers as GPT, but which has been trained on a larger and better data set (and with other changes), making it remarkably good at generating (near) human-quality fake text.  
%
By comparing the poorly-trained (i.e., ``bad'') GPT to the well-trained (i.e., ``good'') GPT2-small, we can identify empirical indicators for when a model has in fact been poorly-trained and thus may perform poorly when deployed.
By comparing GPT2-medium to GPT2-large to GPT2-xl, we can examine the effect of increasing data set and model size simultaneously, as well as analyze well-trained versus \nred{very-well-trained} models.

@@

We now look across series of increasingly improving GPT2 models (i.e., we consider well-trained versus \nred{very--well-trained} questions), by examining both the PL exponent $\alpha$ as well as the Log Norm metrics.  
In general, as we move from GPT2-medium to GPT2-xl, histograms for both $\alpha$ exponents and the Log Norm metrics downshift from larger to smaller values. 
For example, see Figure \ref{fig:gpt2-histograms}, which shows the histograms over the layer weight matrices for fitted PL exponent ($\alpha$) and the Log Alpha Norm ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$)~metric.

We see that the average $\alpha$ decreases with increasing model size, although the differences are less noticeable between the differing well-trained 
versus \nred{very-well-trained} GTP2 models than between the poorly-trained versus well-trained GPT and GPT2-small models.
Unlike GPT, however, the layer Log Alpha Norms behave more as expected for GPT2 layers, with the larger models consistently having smaller norms. 
Similarly, the Log Spectral Norm also decreases on average with the larger models (not shown).  
As expected, the norm metrics can indeed distinguish among well-trained versus \nred{very-well-trained} models among a series well-trained models.
